\documentclass{article}
\usepackage{parskip}
\usepackage{pdfpages}
\usepackage[margin=.6in]{geometry}
\begin{document}

\section*{The Object}
\label{sec:the_object}
Enola Gay is the plane that dropped the first atomic bomb, named after his mother. Paul Tibbits was the pilot who flew it.

\section*{The Event}
\label{sec:the_event}
Leslie Grows picked tibins to delever the payload. He was told to get the crew together and plan the mission. They start to approach the target and a crew member preps the device. It gets dropped. A lot of people died in the city. Most of them were civilians.

\section*{The Reaction}
\label{sec:the_reaction}
The bomb drop was probably race based. People didnt really know the power of nuclear. Lots of things changed.

\section*{The Debate}
\label{sec:the_debate}
People look at the dropping of the atomic bomb was a choice made by the president. It was a hard choice. The argument was that the dropping of the bomb saved thousands of american troups that would have kept fighting. 500,000 might have been an inflated number from stimson.

Herbert hoover felt the use of the atomic bomb was immoral. Curtis Le May felt that the bomb was not needed to end the war, but his word is suspect since he was incharge of the air attacks on japan and felt that he was forcing the end of the war through conventional weapons.

In 1965 a book is realeased. Its the first substantial argument against the dropping of the bomb. He was heavily influenced by the wars in vietnam and the political environment of the 60's. He argues that this is not the end of the second war, but the start of the cold war. This is so bad because the war was already pretty much done and the bomb was dropped in order to scare the soviets.

Some people think that the bomb had to be dropped because the government had spent that much money on it.

So there is tons of skeptics and pressure from various people to get an exhibit up.

\section*{The Exhibit}
\label{sec:the_exhibit}
The debate is on how to display the enola gay. It was acquired by the smithsonian. The museum didn't really have a place to put it. Its just left out where the weather destroys it and it gets vandalized.

They came up with a solid script for the exhibit. Many of the people who previewed it did not fully read it and took many lines out of context. The exhibit needed to be defended before congress. One of the arguments was that history should be a recitation of facts with no interpretation. The exhibit gets killed, the script it not allowed to be released.



\section*{The Conclusion}
\label{sec:the_conclusion}








\end{document}
