\documentclass{article}
\usepackage{parskip}
\usepackage{pdfpages}
\usepackage[margin=.6in]{geometry}
\begin{document}
At first we thought everything was either a particle or a wave, but then we discovered a third option with properties of both, called a quantum particle/

Light was first considered as a wave, but it requires a medium. Huygens called this \emph{ether} which uniformly fills space (thing of dark matter). Light was also a shower of particles moving through a vaccume (Newton's theory). They argued back and forth for a hundred years but couldnt prove anything. Young proved definitively that light was a wave experimentally. This made all physicists move away from newton's theory.

%21
We use a paricle called a kaon that decays at a steady rate, so we us it as a clock and accelerate it super fast to see how the time it take to decay and see how it is effected.


Yeaaaahhhh really didn't pay attention here, its just a long ramble about time dilation. Weee




%28
Say we have alice movign 1m/s but for bob observing from the sides space-time is being dilated so it seems like alice is moving less than 1 meter and its taking longer time to do so. As speeds increse, so does dilation. The speed of light is limited by that relationship asymptotically.

Say some mass is in alice's ship moving at v. If bob tries to accelerate that mass he has to apply force over time. But bob and alice have different values of time. So to bob the mass will feel like it has more mass.

%39
There is time dilation and space dilation in science. There is also relativity of simultaneity. Say we have a ladder far too big to fit in the garage, if we move the ladder very quickly it will have length contraction so it will fit in the garage. For an instant it will anyway. If we look at things from the point of view of the ladder the garage is moving fast. This means that the garage will length contract making it even smaller, which means that the ladder will not fit.



\end{document}
