\documentclass{article}
\usepackage{parskip}
\usepackage{pdfpages}
\usepackage[margin=.6in]{geometry}
\begin{document}
\title{Crusades}
\maketitle
Keywords:
\begin{itemize}
	\item Alexius Comnenus
	\item Ausustine of Hippo
	\item Agustus, Philip
	\item Barbarosa, Fredrick
	\item Clermont, France
	\item Doctrine of the Treasury of the Church
	\item Franciscan and Dominican friars
	\item Indulgences
	\item Jerusalem, Conquest of
	\item Just Cause
	\item Manzikert, Battle of
	\item Outremer
	\begin{itemize}
		\item Antioch, Principality of
		\item Edessa, County of
		\item Jerusalem, Kingdom of
		\item Tripoli, Country of
	\end{itemize}
	\item Papal legates
	\item Peasent's Crusades
	\item Piacenza, Council of (1095)r
	\item Redemption of indulgence
	\item Richard I (the lionhearted) - King of England
	\item Salah ed-Din, or Saladin (1138-1193)
	\item Stephen and Nicholas, Children's Crusade (1212)
	\item Pope Urban II
	\item Zengi, 24 Dec 1144
\end{itemize}

\section*{Basic Timeline}
\label{sec:basic_timeline}
\begin{itemize}
	\item -1000: Egyptians, ish
	\item -500: Roman Empire
	\item 500: Fall of Roman Empire
	\item 1500: Renaissance
	\item 1500-Current: Modern History
	\item 500-1500: Middle Ages
	\item start-500: Ancient History
	\begin{itemize}
		\item Early Middle age (used to be called the Dark ages)
		\item High Middle age
		\item Late Middle age
	\end{itemize}
	\item 1250: Global cool off causes environmental disaster resulting in 7 years of famine (the people that survived this were hella fucked healthwise)
	\item 1300: Black Death (about half the world dies) which lead to people starting to question the church
	\item early 1300: protestant reformation
\end{itemize}

November 27, 1095 Pope Urban II started the crusade to reclaim the holy land from the turks. This kicks off a thousand years of holy wars against muslims all over the known world.

\section*{The Events}
\label{sec:the_events}
Pope Urban came from a rich family. Worked towards getting rid of corruption in the church. He was a monk which reject the world (promoting the religion by being a role model). He puts in the rule that priests could not be married. People did not like this because it makes preists lest approachable (cannot bring marital problems to them). This stated Gregorian Reforms, it also starts the bickering between the church and the government. This eventually lead to the separation of church and state which caused the government to invest in banking and education which causes a boom of innovation and lets the western world jump ahead.

Battle of Manzakert 1071's Turks are taking over the remnants of the eastern roman empire (Byzantines). They have a conclusive battle and capture the emperor. Theres a but of emperial struggle in the turks. This causes the people of constantinople to send a request for help to the church. This lead the pope to make this first call for help from christandom.

The leader in Constantinople wanted knights to fight and the Pope wanted a political win. Only poor knights and farmers responded to the popes call and marched east without real training, thats why its called the Peasents Crusade. They marched east and slaughtered jewish villiages as they went (did bunch of bad things). When they arrived the turks destroyed them.

The Knights finally got organized and started heading over (this was the offical First Crusade). At the time they arrived the Turkish leadership had died. This resulted in the Turkish Empire being divided so the knights won easily. They took the holy land and made four new states collectively called outremer. This was the first time in a while that the christian kingdoms had to figure out how to collonize and spread their culture. Alot of the money and myths about "natives" that went with the exploerers of the renaissance come from this.

Eventually the muslim world reunited and decided that the best way to stabalize their empire was to kick the french out of outremer. A charismatic muslim leader Zengi conquered Edessa on christmas eve. This lead europeans to believe they had lost the grace of god. In response to this they sent a second crusade which failed miserably.

Islamic leadership was taken over by a well educated Turkish leader named Saladin. The Battle of Horns at Hattin (1186) was the crushing defeat that ended it. This is also the location where the Jesus gave the Sermon of the Mount. This also leads Europeans to believe they had lost the grace of god.

Saladin captures basically everything except for Tyre. They ask for help which triggers the Third Crusade. Richard the Lionheard (England), Barbarosa (German), Phillip II. They don't really do well.

The fourth crusade was by Pope Innocent III. The pope lost control of the forces and they got diverted by spies resulting in them sacking constantinople, the city that asked for help in the first case. This caused a ton of resentment.

1291 was basically the end of all the crusading shenanigans.

\section*{The Context}
\label{sec:the_context}

\subsection*{Just War}
\label{sub:just_war}
Crusading is rooted in very old religious ideas (ideas older that Christianity), from Judaism. The 10 Commandments were the most important rules to Judaism where the 5/6 is no killing. Different sects count the commandmants differently due to grammar. Early christian communities followed this commandment quite religiously. Augustine of Hippo (354-430) wrote two books (the confessions, and the city of god) where he tried to understand by bad things happen to good people. Barbarians were at the gates of his city. The material city is where bad things happen but by living the good life you can get to the heavenly city where bad things do not happen. He also wrestles with the problem of not killing so he comes up with some rules about when killing is ok.
\begin{itemize}
	\item just cause (response to aggression)
	\item Sanctioned by prince (or anyone with power)
	\item right intention (whats in your heart at the time)
\end{itemize}

For mideval thinkers the act of violence was a neutral act, but it was based on these rules.

\paragraph{Des Levot}
\label{par:des_levot}
was the battle cry of the crusades. It means God wants it.

As people traveled they stated to incorporate other ideas into crusading. One of these was Penance.

\subsection*{Penance}
\label{sub:penance}
A priest can cleanse your soul of sin by having you do something good. One of these things was going on the crusades. It could be considered a prilgrimage to the holy land. They left fancy clothes behind and didnt wear shoes. They were also meant to fast before battles. Basically you should be reminded that this is painful. The pope grants prilgrim status to all crusaders and had them take a vow (there were massive fancy ceromonies) where they got a red cross to their clothes which they could not take off until they got back. The vow made them temporary members of the church, giving them special rights
\begin{itemize}
	\item could not be prosecuted
	\item their lands were protected
	\item could not be taxed or tolled (a big deal since mobility was rare)
	\item could get back from being excommunicated
	\item the vow substituted any other vow
 \end{itemize}

By making it a holy war they could not limit who signed up so lots of undesirables tagged along. There were tons of people following along to do business with them. They church had to keep setting laws about it (like no whores on crusade). The army also moved very slowly because of all the undesirables. The church also incentivized it more. They would give people indulgences for listening to the "sales pitch" for the crusades. These weren't full indulgences (where everything was forgiven), but just shortened the time you spent in purgatory. Bernard Clairvoux was a popular preacher that would promise full indulgences.

The churched claimed that saints had lots of "grace" which the church collected when a saint died. This meant that the church had infinite grace that they could give out. This was called the \textbf{Treasury of the Church}. People also believed martyrs were gaurenteed entrance to heaven.

\section*{Marketing and Mounting a Crusade}
\label{sec:marketing_and_mounting_a_crusade}
By the time of Innocence III they were signing up anyone that showed. There were also a tone of preachers that went off message

Raised a ragtag army through evangelization to go on the crusades, they all died.

Some boy preachers Stephen and Nicolas raised children's armies. This lead to the myth of the Pied Piper. Charismatic preachers would raise armies of children to go fight but they were actually sold into slavery.

The cost of raising and transporting armies became too expensive which lead to the end of things. The Pope was dependant on partnership arrangements (needed support of kings/governments). There were times that France was spending 60 times the yearly income on a crusade. The pope could also tithe and tax. There was a while where all fines for blasphemy went to the crusades.

People would often leave money to the crusades in their will.

The church gave people the option to quite their vow to the crusades by \textbf{Redeeming} their indulgence by paying the equivalent cost of going on the crusades.

The pope was technically the leader of the crusades but they are not allowed to fight so they delegate to others which cause problems.

The basic theory is that people aggreed to go on the crusades out of religious devotion.


\section*{Summary}
\label{sec:summary}
Crusading forced europeans at every level of society to work together and learn the logistics of large scale war.

Crusading brought christianity into direct contact with islam. The islamic nations were much more advanced in knowledge which made it to Europe when they bring back books from musim nations. Muslims had preserved Aristotle's texts which the church had lost.

The disaster of the fourth crusade wided the rift between west and east church.

Learning about colonization.

Religious innovation.

Created the notion of religious wars.

































\end{document}
