\documentclass[12pt]{article}
\usepackage{parskip}
\usepackage[margin=.2in]{geometry}
\title{Psychopathy: Schizophrenia}
\begin{document}
\section*{History}
Eugen Bleuler was the first to really document this disorder (was the man who critized Binet). He coined the term schiz (split) phren (mind) to discribe a contrast between two minds that shouldnt be there. A split in the ABC's of the mind.
\section*{Myths}
\begin{itemize}
\item this is not spit personality disorder
\item schizophrenic individuals are dangerous - their behavior is very bizzare and inappropriate, Not Intentionally Harmful. 
\item its a permanent condition - it is usually a life long battle, we tend to not want to declare them recovered (incase they do something), but many make a good recovery
\item schizophrenics are creative - not necissarily
\end{itemize}
\section*{Diagnosis}
Based on DSM-IV criteria:
\begin{itemize}
\item must have two  symptoms of :
\begin{itemize}
\item missed these, get them from slides later
\end{itemize}
\item social/occupational dysfunction: on or more major areas of function are markedly below previous level
\item duration: continuous duration for at least 6 months
\end{itemize}

\section*{Symptoms}
\begin{itemize}
\item positive
\begin{itemize}
\item hallucinations (things not based on anything in this world)
\item delusions (misunderstandings based in reality) especially persecutory
\item disorganized thought and nonsensical speech
\item bizarre behaviors
\end{itemize}
\item negative
\begin{itemize}
\item flat affect - no emotion in face
\item no social interaction
\item anhedonia - no felings
\item acolition - no motivation
\item alogia - no speaking
\item catatonia - no movement
\end{itemize}
\end{itemize}

\paragraph*{Problems in Thinking and Speaking}
People with schizophrenia have great problems thinking, often they even have something called word salad where their sentences are completely nonsensical. Delusions are very common, especially in paranoid. Delusions are usually a very complex and set belief usually of grandeur or persecution. Messianic is the delusion that you are or were chosen by a god. They also have problems with selective attention, cant chose which thoughts to act on or say out loud. 
\paragraph*{Disturbed Perceptions}
Hallucinations are a common symptom, the most common of which is hearing voices, rarely they might see, smell, touch, or taste weird things that arent there. 
\paragraph*{Inappropriate Emotions and Behavior}
Not responding to the environment, or responding in compeletely the wrong manner. Often have errors in social perception, disorganized, unfiltered thinking, delusions and hallucinations. Sometimes the body exhibits symptoms repetitive or even catatonic where they freeze motionless randomly.
\section*{Subtypes}
\begin{itemize}
\item paranoid - hallucinations with negative messages and delusions
\item disorganized - flat affect (no emotion), incoherent speech, random behavior
\item catatonic - rarely initiating or controlling movement; copies others' speech and actions
\item undifferentiated - everything else
\item residual - people in withdrawl
\end{itemize}


\section*{Causes}
There is decent amount of evidence for a genetic factor to the cause of scizophrenia. The evidence isnt completely successful, mearly just studies on twins and families and such. Physically we know that schizophrenia is directly related to excess levels of dopamine. Similar symptoms to an amphetamine overdose all the time. Poor firing in the frontal lobe causing reduced judgement but lots of activity in the thalamus acts just like the hallucinations are real. General shrinking of the braina nd enlarging of the ventricals, but this might be a result of schizophrenia. Psychologically not much causes schizophrenia, just the affect of onset. Some stressors that might bring about the onset of schizophrenia might be economic class (lower class is much more prevalent). A schizophragenic family is a family that makes the symptoms worse (having multiple conversations at once, discounting the son's abilities, ignoring him, etc).

\section*{Therapy}
 Pharmacological treatment - antipsychotics - chlorpromazine was developed in france a a cure for the cold (antihistamine) and they tested it on mental health patients. It binds to the dopamine receptors and acts as a blocker. Too much of it can result in parkinson symptoms or Tardive Dyskinesia (result of long term use, no fix for it) which is a continuous need to twitch. Schizophrenia has a high remission rate. Some people dont like the side effects of the medication and some start to think that they were the cause of all improvements and they stop taking the medications. Sometimes the cause for remission is that the end up back in the same shitty environement, the world is not kind to them.
 \paragraph*{Milieu Therapy} Created at McMaster. To be treated you must be recovering from a psychotic episode. In this therapy they had a small economy, had jobs and had to pay for things, the difficulty and rights would increase over time as they got better. This prepared them much better for the outside world.  













\end{document}