\documentclass{article}
\usepackage{parskip}
\usepackage{pdfpages}
\usepackage[margin=.6in]{geometry}
\begin{document}
Basically this is the time when the world said we should give rights to indigenous people. Took 20 years to make. Consitutes the minimum standards for the survival of natives. Protects the collective rights that may not have been addressed by other rights declarations that focus on individual rights.

\paragraph{who are natives}
\label{par:who_are_natives}
People must self identify as native, acceptance by community,. Historical community pre colonial. Strong link to territories and natural resources, distinct social systems, distinct culture, non-dominant groups of society, resolve to maintain traditions.

\paragraph{list of rights}
\label{par:list_of_rights}
foundation rights, life and securty, language and culture idnetity, eduction and social services, percitimation in government, rights to country rules, etc

Resident school were no bueno. The indian act was a policy that made natives wards of the state of canada, it treated them like children. Its main cause was that the government thought it needed to take care of them. Frequently it was used to try to convert natives into ``good christians''. The main goal was assimilation. The education was very poor, children only learned for half the day and worked the rest of the day to provide the labor to fund the school. These schools lead the the right to not be assimilated into the charter of rights.

Health issues were a big deal at the schools. This was handled by discharging the sick students and sending them home. A big problem was people being removed from their homes so article 10 was a big deal for them.

In the marshal islands the us tested a ton of nuclear weapons. They had to move all the native people there and they did so without consent or payment. The government also wanted to know about the effects of radiation on humans so they tested on the people on the islands. They claimed that a shift in wind brought the radiation to the island they lived on.

There was a case of canada moving nine families to the artic and did not allow them to return.






\end{document}
