\documentclass{article}
\usepackage{parskip}
\usepackage{pdfpages}
\usepackage[margin=.6in]{geometry}
\begin{document}
Many of the resources we have about the erradication focus mostly on what happened while we were trying to get rid of it. More modern resources complicate things due to the politics of the cold war. Due to the cold war the USA and Russia would combat each other by trying to influence smaller countries. These countries would try to play them off each other to get help, but this help often came with other consequences.

Small pox has two forms, major and minor. It ws not a fun disease. Variola major killed about 25\% of people that caught it, the minor version only gets about 1/2\%. The disease is air born, you only really need to be within 7ft of a person with it. When it enters it goes straight for the lymph nodes to get your body to replicate it. After around 14 days you start to see symptoms. A very high fever, head ache, and back ache. These are not weird symptoms. The weird symptoms are the spots that appear and spread all over your body. It starts as a flat red rash and turns into raised bumps (these signify the second stage). The bumps are filled with pus and extremely painful. They eventually hardden. Eventually they fall off and turn into pock marks.

This was a target for eradication because there are no animal vectors or resiviors. This means that you only have to make all humans immune to kill it. They saw that most people with the minor version survived and became immune, unfortunatly the major one was the dominant one.

The main way to deal with it was to cut into the pustules and taking materials and putting it into other people so that they would get a mild case. The chineese and indian subcontenent took the pustules that fell off and ground it up into a nasal spray. These methods very rarely worked and often just spread the disease.

At the end of the 18th century a doctor known as edward jenner had been innoculated and nearly died. There was a myth that mild maids that got cow pox did not catch small pox. Small pox came in waves every 5-10 years but these milk maids didn't catch it. Jenner took material from the cow pox pustules and used that to innoculate a young man. He then innoculated that boy with small pox, and he didn't catch it. He tried again a few times and it all worked so he submitted his research and it was published by the Royal Society in 1796.

They felt that by using the cow pox it wouldnt contain the deadly version of the small pox. When they announced this result some private places tried to do their own version which was a mix of cow and human small pox which didnt work at all. This caused people to be skeptical.

This was the difference between vaccine and inoculation.

Before the cow pox version skeptisism about innoculation was pretty valid, most of the time these things did not work or even made things worse. This required very very special preparation and would prevent them from going to work. The working class was very opposed because they couldnt afford to not go to work and it often killed their children. They did not want mandatory vaccination.

A lab found that you could purify the vaccine with glicerine. The vaccine also had to be kept cold. It wasnt until the centerfuge that we found we could freeze dry the vaccine.

It was not uncommon for outbreaks to happen when people travel from other places where the disease was still prevalent. This caused people to want to erradicate this across the whole world.

The world was already becomming a more unified place. We also were already focusing on good health. This lead to the creation of the WHO. This was based on the efforts of the league of nation's health organization. The united states and russia had not belonged to the league of nations so there was some problems. They wanted everyone to be able to join and wanted perspective from a global standpoint, not a national one. They felt national politics had too big of a role in health endeavors.

In order to join you had to pay an annual fee that was based on the gdp of the country. Politics of course was going to fuck this up. The usa was going to be a leader of the UN and the head quarters was going to be in new york. The devloping countries were not ok with this so it was put in Geneva. The director of the WHO was elected through a competition and the winner was a canadian dude. This guy eventually retired because he could not stand the politics.

The guy who replaces him is a brazillian. Brazil had some of the highest small pox in the americas. He wanted the erradication but felt that it was not possible. Problems also occur when russia pulls out. After stalin dies russia want to start interacting with the world, they send a envoy with Viktor Zhdanor that brings up the notion of erradicating small pox. He makes the point that erradicating it in your country is not enough as you have lots of people trying to enter all the time. Russia promises tones of vaccines if the WHO would create a vaccination program. The americans, wanted to have a malaria erradication campeign and felt that the russians were insulting them. The WHO does not have the funding to do both projects. So they set up a tiny small pox team and give them a very small amount of money and ingnore it.

Things change when the first case in a while occurs in the us when missionaries bring back the disease from afar. Fortunatly they caught this and quarenteened the family. The ontario government did not want to do a mass immunization campaign in response to this as the vaccine needed to be refined to not do damage to people. There were a couple of other small out breaks which lead to the idea that small pox can travel on air planes. Instead of having to organize a mass immunization campaign we should go to where the disease exists and nip it in the bud.

Lyndon Johnson toured india and saw  the poverty and poor health of the country which lead him to lend support for the erradicate the small pox. He likened it to putting a man on the moon. The CDC at the time was competing for congressional funding so they jumped on board with the small pox vaccine.

They believed in shoe leather aid, if requested they would send a team to the locaiton of an outbreak to look for the source.

Basically a whole bunch of politics went down. People thought it would fail and wanted to make sure they could pin it on the americans. Eventually we got some people in charge that actually wanted to help people.

A nonprofit organization developes the freeze dried vaccine and they are willing to share it in the world. So yay, we got a good vaccine.

Not we are faced with the problem of how to vaccinate people in hard to reach places. The americans invented a gun that could do a bunch of vaccines very quickly. This ran into problems in unhygienic conditions which clogged them up. They then invented a biforcated needle that was specially design to just dip and poke people. This couldnt get clogged up as easily. This was almost as fast as the gun and was way lest frightening. It was also very easy to use, so people could train locals on how to use it so that the people could get their vaccines from people they trusted. Many people were scared of foreigners.

They realized they couldnt vaccinnate everyone and instead jut vaccinated those around people that had small pox. We set the guide line that a country is case free for two years then we can say it is small pox free.

TADA its done, we killed it. Yay. What do we do with the remaining stock of vaccines. Two russian scientists deffected and told stories about russia weaponizing diseases. This lead people to be concerned that maybe there are more vials of the virus.

People were worried about terrorism and the WHO was given permission of have some of the small pox virus just in case. In the process they found a freezer with some left over vaccine. Other places have found other unknown small pox virus vials. Basically we have no idea where all of the small pox stuff is. It is possible that this could be use in bioterrorism. You need a sophisticated labs and equipment to do it. Basically we are not to worried about it, we are pretty good at dealing with diseases. Just a few days ago we invented a new vaccine for small pox just incase.















\end{document}

