\documentclass{article}
\usepackage{parskip}
\usepackage{pdfpages}
\usepackage[margin=.6in]{geometry}
\begin{document}
\title{Crusades}
\maketitle

\section*{History}
\label{sec:history}

The first printed works, mainly religious in the 1454-1455 in Gutenberg, Germany. The gutenberg bible was the first major book produced using metal type. Its 1286 pages most on handmade paper. Before this monestaries had scriptoriums where monks would copy things out by hand. We now understand that most of these monks were illiterate and were just replicating the shapes they saw. They were also told to not correct mistakes. This resulted in books being very expensive and full of errors.

Movable type was actually first invented by the chinese in 1040's and it spread to korea as well. They used clay and wood for their tiles. It caught on way more in europe due to the simplicity of the alphabet.

Gutenberg invented oil based ink because water based ink would not stick to the metal. There was also a huge explosion of paper making, which made it much more affordable. Paper also got started in china.

\section*{Effects}
\label{sec:effects}

They spread all over europe. An estimate, before printing press 30,000 books in europe, after 10-12 million.

This also standardizes books as well as making them affordable. You cannot share science if a clerical error can occur.

We also introduce the concept of owning knowedge. This didn't matter when there were so few books. This results in copyright

More people can own books. Only the church had access to the bible so they could interpret it for you, but now you can own a bible which shifts power. This has a trickle down effect of encouraging protestantism. This also results in the rise of censorship since knowledge is becoming widespread. The church decides that nothing could be printed without the permission of the church. The government also does this with king charles of france.

This also revolutionizes scholarship as people can publish their own works which means that scholars can spread their discoveries much more quickly and reliably. The stability of the data in their works is reliable. Great scientists emerge in this period due to this including galileo, copernicus, decartes to name a few.

Society now has the ability to remember much more and more reliably.

The media ecosystem expands, we have a new way to communicate. Rich people can reach larger audiences. Printing becomes cheaper so we can spread more and more knowledge.

A limit to this is that only the rich can aford to have their works published.

\section*{Internet}
\label{sec:internet}
Very likened to the printing press.
