\documentclass{article}
\usepackage{parskip}
\usepackage{pdfpages}
\usepackage[margin=.6in]{geometry}
\begin{document}
Basically we spent a bunch of time looking for a biological cause of homosexuality. This one guy found that a certain brain formation may be bigger or smaller. This study was a bit inconclusive as the sample size was quite small and they could not reproduce it. They did find similar patterns in sheep though.

Another study finds that men with more older brothers are more likely to be gay. They think this might be the mother's immune system. The mother's immune system develops antibodies when she has sons and with each new son more of those antibodies get into the embryo which alters the sexual orientation.

Foucault wrote the \emph{History of Sexuality} which brought much more attention to this. Previously homosexuality was a sin, then a sickness. We first wanted to punish it, then to fix it. Now that we are working to a medical model we are looking to understand it, or maybe ``cure'' it.

Sodomy represents what I did, practice. Homosexuality represents who I am, personality.

We have some very basic dualism between sex and gender. We didn't always see the distinction. The second wave feminist movement 1960s-1970s saw sex as the biological basis that gender could be ascribed. Sex is biology and gender is a societal concept. The third wave movement started to question this.

Historically we thought that there was one sex that just appeared differently. Basically we had all the same bits just some flipped inside out. During the enlightenment we started to have two genders and drawing the opposites argument.

\paragraph{John Money}
\label{par:john_money}
Honed the concept of gender in the context of 1950's. He wanted to handle infants with ambiguous genitalia. He is famous for a disastrous experiment. He felt gender was a social construct and its dominant, you could override biology. A boy who's penis was burnt off was raised as a girl. He did not like being a girl, he felt that he was a boy but his parents forced it on him. Eventually he committed suicide.

For a while we dissaproved of homosexuality based on abrahamic religion. As science got better medicine took dominance since they had more knowledge to use. Medicine also provided a more intimate proximity to the patient to get a more honest narrative. Also medicine was affiliated with science and healing so they are known as being compassionate and rational.

\paragraph{Etiological Frameworks}
\label{par:etiological_frameworks}
in 1860s western european physicians commmented on the constitutional basis of homoexuality
\begin{itemize}
	\item naturalists - it is natural phenomenon, a genetic predisposition, it is harmless, homosexuality might just be a third gender
	\item degenerationists - homosexuality was an hereditary defect that caused them to drop on the evolutionary scale, it was a disease and a problem
	\item psychogenists - homosexulaity is psychologically caused, not inborn, it is a perversion of the sex drive from the normal lawsobject to something else
\end{itemize}


\paragraph{Naturalists}
\label{par:naturalists}
\textbf{Kertbeny} argued that there are two germs, one for sexual orientation, and one for sexual drive. \textbf{Hirschfeld} felt that homosexuality was natural like what hand was dominant and did the first big study on it. He also was a activist for it and formed the first organization to try to fight anti-homosexuality. \textbf{Ellis}, \textbf{Carpenter}, and \textbf{Symonds} were british sexologists. They promoted free-love doctrines. Wrote the first medical monograph about sexual inversion. The chinese also had a focus on sexology, they were tyring to do the same thing as the europeans.

\paragraph{Degenerationists}
\label{par:degenerationists}
\textbf{Krafft-Ebing} write the psychopatia sexualis, it was meant only for experts but became quite popular. This is the first time we encounter words for a lot of the categories of sexuality (sadism, masochism, etc). He said homosexuality was like a psychic hermaprhoditism. Since the animals that are hermaphrodites were simpler animals he said that homosexuality was a drop in evolution. He was very negative in his view of homosexuality.

\paragraph{Psychogenists}
\label{par:psychogenists}
Freud thought homosexuality was a perversion of the sex drive. He also thought we were all homosexual an it might appear whenever, called latent homosexuality. Later thinkers thought everyone was born bisexuality which was built on Frueds theory. He felt that the family has an effect on homosexuality. The problematic lifestyle of homosexual could be treated by healthcare professionals (that its their mistreatment that causes sadness, not their condition). Frued was not critical of homosexuality, his American followers were.

\textbf{Hall} spread knowledge on sexologists in his novel \emph{The Well of Loneliness}. It was banned which brought about more popularity. \textbf{Margret Mead} and \textbf{Ruth Benedict} were popular intellectuals that also helped spread knowledge on it. They talked about gender and sexuality and how they were more fluid than we believed.

\textbf{Kinsey} made the united states the center of sexology. He wrote a book \emph{Sexual Behavior of the Female}, also called the Kinsey report. He interviewed all sorts of people about their sexuality. He made his interviewees more comfortable so they could talk more openly. It made psychologists question their beliefs about sexuality.

\textbf{Hooker} published a study that overturned the idea the homosexuality was a disease. People cannot tell the difference between people based on their gender based on her test.

\paragraph{Reverse Discourse}
\label{par:reverse_discourse}
Theory from Foucault. He explained that we need to be careful with the names we use to describe things as they might have effect on how we interpret things.





\end{document}

